\input jtexinfo.tex @c -*- texinfo -*-
@settitle Demacs, 386/486 DOS version of GNU Emacs
@setfilename demacs.info
@iftex
@a4book
@c @smallbook
@c @headings off

@titlepage
@sp 4
@center @titlefont{Using Demacs}
@sp 2
@center A Guide to the 386/486 DOS version of GNU Emacs
@sp 2
@center Version 1.2.0
@center December 12, 1991
@sp 6
@center @secrm{Manabu Higshida}
@center @t{manabu@@sigmath.osaka-u.ac.jp}
@center Osaka University, JAPAN
@sp 2
@center and
@sp 2
@center {@secrm HIRANO Satoshi}
@center @t{hirano@@tkl.iis.u-tokyo.ac.jp}
@center University of Tokyo, Japan
@page

@rightline {Special Thanks to}
@sp 1
@rightline {Morgan Schweers}
@rightline {Shigeki Yoshida}
@rightline {Junn Ohta}
@rightline {Toshihiko Shimokawa}
@rightline {Makoto Matsushita}
@sp 1
@vskip 0pt plus 1filll
@leftline{Copyright @copyright{} 1991 HIRANO Satoshi, Halca Computer Science Laboratory}
@leftline{Copyright @copyright{} 1991 Manabu Higashida}

Demacs is free software; you can redistribute it and/or modify it under
the terms of the GNU General Public License version 1 as published by
the Free Software Foundation.

This program is distributed in the hope that it will be useful, but
WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
General Public License (this package includes it as file @file{copying})
for more details.
@end titlepage
@page
@end iftex


@node Top, Overview, (dir), (dir)

@menu
* Overview::     
* Installation:: 
* Invocation::   
* Features::      
* Restrictions:: 
* Compilation::  
* Appendix::     

These nodes contains followings.

Overview:
* Version::                Current Version.
* Platform::               Platforms.
* BugReport::              Bug Report.

Installation:
* FileList::               Contents of This Distribution Kit.
* InstallationProcedure::  How to Install.
* InstallationExample::    Sample Session to Install.

Invocation:
* EnvironmentVariable::    Environment Variables.
* OptionsOfdemacs.exe::    Options of DOS Extender demacs.exe.
* OptionsOfDemacs::        Options of Demacs.

Features:
* FileType::               File Type: Text or Binary Translation.
* Pass8::                  "8bit clean" Display Mode.
* int86::                  8086 Software Interrupt.
* MachineType::            Machine Specified Features.
* OtherFeatures::          Other Additional Demacs Features.

Appendix:
* TerminalFeatures::       Terminal Specified Features.
* DiffsFromUNIX::          Differences from UNIX Version.
* Index::                  Index: Functions and Variables.
@end menu


@node Overview, Version, Top, (dir)
@chapter Demacs version 1.2.0

Demacs---386/486 DOS version of GNU Emacs.

@menu
* Version::    Current Version.
* Platform::   Platforms.
* BugReport::  Bug Report.
@end menu


@node Version, Platform, Overview, Overview
@section Current Version

Demacs is currently based on GNU Emacs version @b{18.55} (@b{partly 18.57}).

Demacs version itself is @w{@b{1.2.0} (@b{1991/12/12})}.

@deffn Command demacs-version
Return string describing the version of Demacs that is running.
@end deffn


@node Platform, BugReport, Version, Overview
@section Platforms

@table @key
@item Hardware
@b{386/486 based DOS machines}---IBM PC and its compatible, AX, J-3100, 
or NEC PC-9801, High resolusion PC-98 and its compatible EPSON PC-386

@item Software
@b{MSDOS 3.0 or later}

@b{XMS} manager (@file{HIMEM.SYS}) and @b{VCPI} memory manager
(@file{EMM386.EXE}, @file{QEMM}, etc.) compatible.@refill

Demacs does not support @b{DPMI}, thus it can not be used with extended
mode on Windows.@refill
@end table


@node BugReport, Installation, Platform, Overview
@section Bug Report

When you have troubles with using Demacs, please follow the following
steps before inquire us.@refill

First, read this README and the online-manuals (with Info Browser) to
find answers. Next, consult to Emacs experts around you or on the
networks. In case you decide to send a mail to the authors, you must
read "Bugs" section in the Emacs manual before sending the mail.@refill

Suggestions and bug fixes are welcome with next address.@refill

@example
demacs-bugs@@sigmath.osaka-u.ac.jp
@end example

If you want to know current status about Demacs, send empty mail to
next address.

@example
demacs@@sigmath.osaka-u.ac.jp
@end example

@noindent
We will reply your mail with latest information, automatically.


@node Installation, FileList, BugReport, Top
@chapter Installation

@menu
* FileList::               Contents of This Distribution Kit.
* InstallationProcedure::  How to Install.
* InstallationExample::    Sample Session to Install.
@end menu


@node FileList, InstallationProcedure, Installation, Installation
@section Contents of This Distribution Kit

The package consists of two archives; @file{dem120e.lzh} and
@file{dem120s.lzh}. You need @b{LHarc} or @b{LHa} utility to unpack
these archives. Don't forget to use @code{/mx} switches with @code{-x}
option to unpack them. For example do as following.@refill

@example
lharc x /mx dem120e.lzh
lharc x /mx dem120s.lzh
@end example

@noindent
And similarly apply these options to the archives which are contained by
them.

After unpacking, following files will appear.

@file{dem120e.lzh} (executables)

@ifinfo
@display
readme            this file
demacs.tex        texinfo version of @file{readme}
demacs.dvi        compiled @file{demacs.tex} by TeX
copying           GNU General Public Licence
copying.dj        DJGCC copyright notice
termcap           sample termcap file
_emacs            sample startup file
bin\demacspc.exe  DOS extender for IBM PC and its compatible machines
bin\demacs98.exe  DOS extender for PC-9801, PC-386 series
bin\demacs        demacs object file (stripped)
bin\temacs        `plain' demacs object file (stripped)
etc.lzh           emacs documentation string file and etc.
lisp.lzh          includes lisp files which are necessary to run
edired.lzh        enhanced dired-mode
@end display
@end ifinfo

@tex
% \vskip \baselineskip
\halign{\qquad \tt #\hfil &\qquad #&\rm #\hfil\cr
readme            &&this file\cr
demacs.tex        &&texinfo version of {\tt readme}\cr
demacs.dvi        &&compiled {\tt demacs.tex} by \TeX\cr
copying           &&GNU General Public Licence\cr
copying.dj        &&DJGCC copyright notice\cr
termcap           &&sample termcap file\cr
\_emacs           &&sample startup file\cr
bin$\backslash$demacspc.exe  &&DOS extender for IBM PC and its compatible machines\cr
bin$\backslash$demacs98.exe  &&DOS extender for PC-9801, PC-386 series\cr
bin$\backslash$demacs        &&demacs object file (stripped)\cr
bin$\backslash$temacs        &&`plain' demacs object file (stripped)\cr
etc.lzh          &&emacs documentation string file and etc.\cr
lisp.lzh         &&includes lisp files which are necessary to run\cr
edired.lzh       &&enhanced dired-mode\cr
}
@end tex


@file{dem120s.lzh}  (source files)

@ifinfo
@display
src\diffs.lzh       includes only different files from the original
lisp\diffs.lzh      includes only different files from the original 
go32_98\source.lzh  PC-9801 version go32 DOS extender from DJGCC 
go32_pc\source.lzh  IBM PC version go32 DOS extender from DJGCC    
@end display
@end ifinfo

@tex
% \vskip \baselineskip
\halign{\qquad \tt #\hfil &\qquad #&\rm #\hfil\cr
src$\backslash$diffs.lzh       &&includes only different files from the original\cr
lisp$\backslash$diffs.lzh      &&includes only different files from the original\cr
go32\_98$\backslash$source.lzh  &&PC-9801 version go32 DOS extender from DJGCC\cr
go32\_pc$\backslash$source.lzh  &&IBM PC version go32 DOS extender from DJGCC\cr 
}
@end tex

@noindent
NOTE: 

This package only includes lisp files which are modified for Demacs.
Original and/or other useful lisp stuff can be obtained from various FTP
sites (for example @code{ftp.sigmath.osaka-u.ac.jp}) or from:@refill

@display
Free Software Foundation, Inc.
675 Mass Ave, Cambridge, MA 02139, USA
@end display


@node InstallationProcedure, InstallationExample, FileList, Installation
@section How to Install

Demacs requires special version of @file{go32.exe} DOS extender. This
package includes it with its name @file{bin\demacspc.exe} for IBM PC or
@file{bin\demacs98.exe} for PC-9801 series. You should use one of them
with renameing to @file{demacs.exe}.@refill

Since @code{go32.exe} which came along with the package of DJ's GCC/G++
version 1.05 can not handle @kbd{Ctrl-C} properlly, we modified it to
work correctly. And we made some changes to be able to handle more DOS
function call, and enhanced its memory management strategy (detail, see
below).@refill

Installation procedure is quite easy. 

@enumerate
@item
If your system is IBM PC variant (including AX and J-3100 series), copy
@file{bin/demacspc.exe} to the directory in @var{PATH} with its name
@file{demacs.exe}.@refill

If your system is NEC PC-9801 or EPSON PC-386 series, copy
@file{bin/demacs98.exe} to the directory in @var{PATH} with its name
@file{demacs.exe}.@refill

@item
Copy @file{bin\demacs} in the same directory with @file{demacs.exe}.

@item
Create @file{lisp} and @file{etc} directory in @file{\lib\emacs}
directory.@refill

@item
If you want to use lisp library, get original lisp files from somewhere
and copy them in @file{\lib\emacs\lisp} directory. Then unpack
@file{lisp.lzh} in @file{\lib\emacs} directory.@refill

@item
Unpack @file{etc.lzh} in @file{\lib\emacs} directory.

@item
Set @var{TERM} environment variable for your system (See following example).

@item
Set @var{TERMCAP} variable if @file{termcap} file is not stay under
@file{\etc}.@refill

@item
Prepare your @file{_emacs} (@file{.emacs} on Unix system) under your
@var{HOME} directory if you need it.@refill
@end enumerate


@node InstallationExample, Invocation, InstallationProcedure, Installation
@section Sample Session to Install

@example
C> copy bin\demacspc.exe \tools\bin\demacs.exe (IBM PC, AX, J-3100)
C> copy bin\demacs \tools\bin
     or
C> copy bin\demacs98.exe \tools\bin\demacs.exe (PC-9801, PC-386)
C> copy bin\demacs \tools\bin

C> mkdir \lib\emacs
C> mkdir \lib\emacs\lisp
C> mkdir \lib\emacs\etc
C> copy lisp.lzh \lib\emacs
C> copy etc.lzh \lib\emacs
C> cd \lib\emacs
C> lharc x /mx lisp.lzh
C> lharc x /mx etc.lzh
C> del lisp.lzh
C> del etc.lzh

C> set TERM=ibmpc  (IBM PC)
      or
   set TERM=j3100  (J-3100)
      or
   set TERM=pc98   (PC-9801, PC-386)

C> set TERMCAP=c:\etc\termcap   (IBM PC, J-3100)
     or
   set TERMCAP=a:\etc\termcap   (PC-9801, PC-386)

C> set HOME=c:\home (your startup file '_emacs' is placed)
@end example


@node Invocation, EnvironmentVariable, InstallationExample, Top
@chapter Invocation

@menu
* EnvironmentVariable::  Environment Variables.
* OptionsOfdemacs.exe::  Options of DOS Extender demacs.exe.
* OptionsOfDemacs::      Options of Demacs.
@end menu


@node EnvironmentVariable, OptionsOfdemacs.exe, Invocation, Invocation
@section Environment Variables

@table @code
@item C>set GO32TMP=c:/tmp
directory where swapped files are placed (see DJGCC's readme).

@item C>set TERM=ibmpc
entry name of @file{termcap}.

@item C>set TERMCAP=c:/etc/termcap
directory where @file{termcap} file are placed.

@item C>set HOME=c:/home
directory where startup file @file{_emacs} are placed.

@item C>set TZ=JST-9
time zone.

@item C>set USER=manabu
your name up to 8 characters.

@item C>set EMACSLOADPATH=d:/emacs/lisp
set lisp load path if you don't use standard path @file{/lib/eamcs/lisp}.

@item C>set EMACSEXECPATH=d:/emacs/etc
set etc load path if you don't use standard path @file{/lib/emacs/etc}.

@item C>set DEMACS_KEEP=1024
amount of extended memory to keep (detail, see below).

@item C>set SHELL=c:\command.com
your command interpreter path.
@end table


@node OptionsOfdemacs.exe, OptionsOfDemacs, EnvironmentVariable, Invocation
@section Options of DOS Extender @code{demacs.exe}

@table @code
@item -keep [@var{num}]
When @code{demacs.exe} invokes a child process, @code{demacs.exe} does
not swap out contents of the extended memory to a disk. If @var{num} is
specified, @code{demacs.exe} uses up to @var{num} killo bytes of
extended memory.@refill

You should use this option with XMS or VCPI memory manager. If not, may
cause violation.
@end table

@noindent
More about @code{-keep} opiton:

@code{demacs.exe} provides virtual memory mechanism. If an application
(e.g. Demacs) requires more memory than equipped real memory,
@code{demacs.exe} write out a portion of real memory to a disk in order
to make a free memory, then @code{demacs.exe} assigns the new free
memory to the application. This is called ``swapping''.@refill

Swapping does not only happen when real memory shortage, but also
happens when @code{demacs.exe} calls a child process. That is, before
execution of the child process, @code{demacs.exe} swapps out all of
allocated memory to a disk so that the child process can use memory.
Unfortunatly, swapping takes very long time because Demacs is a huge
program.@refill

In order to eliminate this swapping time, our @code{demacs.exe} can take
@code{-keep} option. @code{-keep} option requests to @code{demacs.exe}
NOT to swap out contents of extended memory (conventional memory (less
than 640KB) is always swapped out).@refill

In case you use @b{XMS} driver such as @file{himem.sys}, since
@code{demacs.exe} obtains all of free extended memory at startup, a
child process can not use any extended memory (Remember GCC/G++ requires
extended memory).  You can reserve extended memory for the child process
by following @var{num} to @code{-keep}, e.g. `@w{@code{-keep
@var{1024}}}'. @var{num} which is killo byte unit number limits amount
of usage of extended memory by @code{demacs.exe}. If less amount of
extended memory is available than @var{num} at startup,
@code{demacs.exe} uses all of extended memory.@refill

Otherwise in case you use @b{VCPI} driver such as @code{QEMM} or
@code{EMM386}, @code{demacs.exe} gets extended memory on demand basis. 
`@w{@code{-keep @var{num}}}' limits amount of extended memory used by
@code{demacs.exe} to @var{num} killo bytes.@refill

The following table shows a brief summary.


@ifinfo
@example
              | swap out          |
              | extended memory   |
              | at fork child     | extended memory for child
--------------+-------------------+------------------------------
without -keep |  YES              | all of extended memory at startup
-keep         |  NO               | nothing
-keep <num>   |  NO               | extended memory at startup - num
@end example
@end ifinfo

@tex
\vskip \baselineskip
\indent\vbox{\offinterlineskip
\halign{\tt #\hfil &\quad\vrule #\quad&\rm #\hfil &\quad\vrule #\quad&\rm #\hfil\cr
            && swap out        && \cr
            && extended memory && \cr
            && at fork child   && extended memory for child\cr
\omit&height2pt&\omit&&\omit\cr
\noalign{\hrule}\cr
\omit&height2pt&\omit&&\omit\cr
without -keep   && YES &&  all of extended memory at startup\cr
-keep           && NO  &&  nothing\cr
-keep {\sl num} && NO  &&  extended memory at startup $-$ {\sl num}\cr
}}
\vskip \baselineskip
@end tex

Too small @var{num} makes Demacs slower. I recommend to set @var{num} at
least 512.

Example (Demacs uses 1024KB extended memory):

@example
demacs.exe -keep 1024
@end example

@var{num} can be also specified by @var{DEMACS_KEEP} environment
variable. If both environment variable and option are specified, option
has priority. If you don't need to specify @var{num} with @code{-keep}
option, specify the value @code{0} with this environment variable.@refill


@node OptionsOfDemacs, Features, OptionsOfdemacs.exe, Invocation
@section Options of Demacs

Type @code{demacs [-keep [@var{num}]] [options]} to invoke Demacs. If
you need more information, see ``GNU Emacs manual''.@refill


@node Features, FileType, OptionsOfDemacs, Top
@chapter Features

@menu
* FileType::       File Type: Text or Binary Translation.
* Pass8::          "8bit clean" Display Mode.
* int86::          8086 Software Interrupt.
* MachineType::    Machine Specified Features.
* OtherFeatures::  Other Additional Demacs Features.
@end menu


@node FileType, TranslateMode, Features, Features
@section File Type: Text or Binary Translation

@menu
* TranslateMode::     Translation Mode.
* BufferManagement::  Buffer Management.
* ReadingFiles::      Reading Files into Demacs Buffers.
* WritingFiles::      Writing Buffers into Files.
* ProcessI/O::        Process I/O.
@end menu


@node TranslateMode, BufferManagement, FileType, FileType
@subsection Translation Mode

On MS-DOS file system, a line is ended with @key{CR} (@t{0x0d})/@key{LF}
(@t{0x0a}) characters and a file is ended with @kbd{Ctrl-Z} character. 
But on UNIX file system, a line is ended with only @key{LF} character,
and end of data meas end of file.@refill

Treating ``binary files'' which contains @kbd{Ctrl-Z} characters and for
compatibility to UNIX file system, C I/O library on MS-DOS has following
two translation mode which control these characters.@refill

@itemize @bullet
@item
Text mode translation
@item
Binary mode translation
@end itemize

With @b{text mode translation}, on reading, @key{CR}/@key{LF} code are
translated to @key{CR} and reading is terminated as soon as @kbd{Ctrl-Z}
appears. On writing, @key{LF} is added to @key{CR} and @kbd{Ctrl-Z} is
added to end of file.@refill

With @b{binary mode translation}, no code is translated.

Demacs supports these two translation mode, and manage them on each
buffer. You can find current translation mode of the buffer from file
type mnemonic on the mode line.@refill

@example
--**-Emacs: demacs.tex   (T:Texinfo)--42%---------------
                          ^
@end example

File type mnemonic means

@table @code
@item T
Text mode translation.
@item B
Binary mode translation.
@end table

We call this translation mode which is managed on each buffer ``@b{file
type}''.@refill


@node BufferManagement, BufferFileType, TranslateMode, FileType
@subsection Demacs Buffer Management

@menu
* BufferFileType::  File Type of Each Buffer.
* BufferCreation::  Buffer Creation.
@end menu


@node BufferFileType, BufferCreation, BufferManagement, BufferManagement
@subsubsection File Type of Each Buffer

File type of each buffer is stored on buffer-local @code{file-type}
variable. To set the value of @code{file-type} use the
@code{set-file-type} function.@refill

The default value of @code{file-type} is the value of
@code{default-file-type} variable. To set its value, use the
@code{set-default-file-type} function.@refill

@defvr {Local Variable} file-type
@table @kbd
@item 0
Text mode translation.
@item 1
Binary mode translation.
@end table
@end defvr

@deffn Command set-file-type TYPE &optional BUFFER
This function sets buffer-local @code{file-type} variable of
@var{BUFFER} to @var{TYPE}. The argument @var{BUFFER} defaults to the
current buffer. The value of @var{TYPE} is one of followings.@refill
@table @asis
@item @kbd{0} or @kbd{'text} or @kbd{"text"}
Specify the buffer's file type to text mode.
@item @kbd{1} or @kbd{'binary} or @kbd{"binary"}
Specify the buffer's file type to binary mode.
@end table
@end deffn

@defvr {Global Variable} default-file-type
The value of this grobal bariable is the default value of buffer-local
@code{file-type} variable.@refill
@end defvr

@deffn Command set-default-file-type TYPE
This function sets the value of @code{default-file-type} variable to
@var{TYPE}. The value of @var{TYPE} is one of followings.@refill
@table @asis
@item @kbd{0} or @kbd{'text} or @kbd{"text"}
Specify the buffer's file type to text mode, by default.
@item @kbd{1} or @kbd{'binary} or @kbd{"binary"}
Specify the buffer's file type to binary mode, by default.
@end table
@end deffn


@node BufferCreation, ReadingFiles, BufferFileType, BufferManagement
@subsubsection Buffer Creation

When creating a new buffer, Demacs set the buffer's local variable
@code{file-type} to the value of the @code{default-file-type}
variable.@refill


@node ReadingFiles, FileInsertion, BufferCreation, FileType
@subsection Reading Files into Demacs Buffers

Using the @code{define-file-name-file-type} function, you can define
file type associated with file name.@refill

For example, 

@example
(define-file-name-file-type "\\.mem$" 'binary)
@end example

@noindent
defines that file type of files having extention @file{.mem} is binary.@refill

By default, the files ending with @file{.elc}, @file{.obj}, @file{.exe},
@file{.com}, @file{.lib}, @file{.sys} (except @file{config.sys}),
@file{.chk}, @file{.o}, @file{.a} and @file{.out} are defined as binary
file type. This is defined in @file{demacs.el}.@refill

@defun define-file-name-file-type FILENAME TYPE
This function defines file type associate with file name. 
@var{FILENAME} is regular expression or @code{nil}. @code{nil}
matches any file anme. @var{TYPE} is file type.@refill
@end defun

@defun find-file-type-from-file-name FILENAME
This function returns file type which associate with @var{FILENAME}
defined by @code{define-file-name-file-type} function. If no file type is
defined, this returns a value of @code{default-file-type}.@refill
@end defun

@menu
* FileInsertion::  Inserting Files into Buffers.
* FileReadingIn::  Visiting Files and Reading.
* FileCreation::   Creating Files and Visiting.
@end menu


@node FileInsertion, FileReadingIn, ReadingFiles, ReadingFiles 
@subsubsection Inserting Files into Buffers

When inserting a file already visiting buffer, Demacs calls the
@code{find-file-type-from-file-name} function with @b{file name of
target file} as its argument, and set return value to file type of the
buffer.@refill


@node FileReadingIn, FileCreation, FileInsertion, ReadingFiles
@subsubsection Visiting Files and Reading

When visiting a file and reading into a buffer, Demacs calls the
@code{find-file-type-from-file-name} function with @b{file name of
target file} as its argument, and set return value to file type of the
buffer.@refill


@node FileCreation, WritingFiles, FileReadingIn,ReadingFiles
@subsubsection Creating Files and Visiting

The @code{find-file-not-found-set-file-type} function is appended to
the @code{find-file-not-found-hooks} variable. This means that
@code{find-file-not-found-set-file-type} is called when the visiting
file not exists.

This function calls internally the @code{find-file-type-from-file-name}
function with @b{target name} as its argument and set return value to
file type of the buffer.@refill


@node WritingFiles, ProcessI/O , FileCreation, FileType
@subsection Writing Buffers into Files

When writing a buffer into a file, Demacs set translation mode to
file type of the buffer.@refill


@node ProcessI/O, Pass8, WritingFiles, FileType
@subsection Process I/O

Demacs communicate with external process (child process) using temporary
file. The @code{default-file-type} variable is used as file type of
the buffer which displays result.@refill


@node Pass8, int86, ProcessI/O, Features
@section ``8bit clean'' Display Mode

Demacs supports so called ``8bit clean'' display mode, which displays
8bit dirty characters as it is instead of backslashed or arrowed
form.@refill

If you want use this mode, execute command @code{toggle-pass8-mode}. You
can find in mode line whether the buffer is in ``8bit clean'' display
mode.@refill

@example
--**-Emacs: foo.bar      (T:Fundamental Pass8)--28%---------------
                                        ^^^^^
@end example

@noindent
To return ordinary display mode, again execute command
@code{toggle-pass8-mode}.@refill

If you want create a buffer always in ``8bit clean'' display mode, place

@example
(setq-default pass8-mode t)
@end example

@noindent
in your startup file.

@defvr {Local Variable} pass8-mode
If this variable is @code{*Non-nil}, Demacs displays 8bit dirty
characters as it is. Default value is @code{nil}.@refill
@end defvr

@deffn Command toggle-pass8-mode
Set/Unset the buffer local variable @code{pass8-mode}.
@end deffn


@node int86, MachineType, Pass8, Features
@section 8086 Software Interrupt

Function @code{int86} generate 8086 software interrupt. DOS Extender
@code{demacs.exe} not handle all interrupts, so some interrupts are
treated as exception and may cause violation. Use this carefully.@refill

Use this function like this.

@example
int
GetDisk ()
@{
  union REGS regs;
  regs.h.ah = 0x19;     /* 25 */
  int86 (0x21 /* 33 */, &regs, &regs);
  return regs.h.al;
@}
@end example

C funtion to get current disk number may write in Demacs as follows:

@example
(defun get-disk ()
  (let ((regs (make-register)))
    (set-register-value regs 'ah 25)    ; 0x19
    (int86 33 regs)                     ; 0x21
    (register-value regs 'al)))
@end example

@defun make-register
Generate instance of register type which is passwd to @code{int86}
function.@refill
@end defun

@defun register-value REGISTER NAME
Get the value of @var{REGISTER}'s @var{NAME}. @var{NAME} is one of
followings.@refill

@table @kbd
@item 'ax
@code{ax} register
@item 'bx
@code{bx} register
@item 'cx
@code{cx} register
@item 'dx
@code{dx} register
@item 'si
@code{si} register
@item 'di
@code{di} register
@item 'cflag
carry flag
@item 'flags
flag register
@end table

@noindent
or

@table @kbd
@item 'al
lower byte of @code{ax} register
@item 'ah
upper byte of @code{ax} register
@item 'bl
lower byte of @code{bx} register
@item 'bh
upper byte of @code{bx} register
@item 'cl
lower byte of @code{cx} register
@item 'ch
upper byte of @code{cx} register
@item 'dl
lower byte of @code{dx} register
@item 'dh
upper byte of @code{dx} register
@end table
@end defun

@defun set-register-value REGISTER NAME VALUE
Set @var{REGISTER}'s @var{NAME} to @var{VALUE}. @var{VALUE} is
@b{unsigned integer}.@refill
@end defun

@defun int86 INTNO REGISTER
Generate 8086 software interrupt of number @var{INTNO} with
@var{REGISTER}. This returns set of register value.@refill
@end defun


@node MachineType, ibmpc, int86, Features
@section Machine Specified Features
	
New lisp variable @code{dos-machine-type} is introduced for support
various machines.@refill

@defvr {Global Variable} dos-machine-type
Set one of @code{ibmpc}, @code{j3100}, @code{pc98} to select machine
specific code.@refill
@end defvr

@menu
* ibmpc::  IBM PC.
* j3100::  Toshiba J-3100.
* pc98::   NEC PC-9801, EPSON PC-386.
@end menu


@node ibmpc, j3100, MachineType, MachineType
@subsection IBM PC

If you use IBM PC, evaluate @code{(setq dos-machine-type 'ibmpc)}, then
@kbd{Ctrl-@key{SPACE}} will generate null code and @code{meta-flag} will
works.@refill


@node j3100, pc98, ibmpc, MachineType
@subsection Toshiba J-3100

If you have J-3100, evaluate @code{(setq dos-machine-type 'j3100)}, then 
following functions will be abailable in addition to IBM PC features.

@example
set-cursor-mode
get-screen-mode
set-cursor-mode
set-keyclick
set-screen-mode
@end example

@noindent
See @file{lisp/dos-fns.el} for more information.


@node pc98, OtherFeatures, j3100, Features
@subsection NEC PC-9801, EPSON PC-386

If you have PC-9801 or EPSON PC-386, evaluate @code{(setq
dos-machine-type 'pc98)}, then following functions will be abailable.

@example
pc98-assign-special-key
pc98-cancel-special-key
@end example

@noindent
See @file{src/dosfns.c} for more information.


@node OtherFeatures, Restrictions, pc98, Features
@section Other Demacs Features

@enumerate
@item
Filename Completion

Demacs provides filename completion feature. Filename may include drive
name in its front. For example, we assume the current directory is
@file{c:/tools/emacs} and there is a file @file{d:/config.sys}, under
this condition @w{@kbd{C-x C-f}} invokes @code{find-file} function then
the function prompts like@refill

@example
Find file: c:/tools/emacs/
@end example

@noindent
at this point your typing of @code{d:/confi} and @key{TAB} key causes:@refill

@example
a:/tools/emacs/d:confi^I  -> d:/config.sys [sole complete]
@end example

@item 
Child Process

Function @code{suspend-emacs} calls sub shell. Try @kbd{C-z} and a new
command shell is invoked.

Function @code{call-process} calls sub shell like @code{suspend-emacs}
but its stdout is redirected to a temporary file and it will be
displayed after completion of the process. Try @kbd{M-!}. But don't call
programs which require input from keyboard, because Demacs can't
response to them.@refill


@item
Enhanced Dired

This package includes enhanced dired-mode by Sebastian Kremer and their
dos ported codes which works without UNIX compatible @file{ls.exe}.

To use this, extract @file{edired.lzh} to your lisp directory and put
lines

@example
(setq dired-mode-map nil)
(load "emacs-19")
(autoload 'dired "dired" nil t)
@end example

to your startup file (@file{direddos.elc} is loaded from @file{dired.elc}). 

If you have @file{ls.exe}, append

@example
(setq dired-load-hook
      '(lambda ()
         (setq dired-ls-program "ls")))
@end example

to your startup file. If you use Demacs with @code{-keep} option, this
configration works faster than former one.@refill

You can get more infomation from @file{dired.README} and its document
string.
@end enumerate


@node Restrictions, Compilation, OtherFeatures, Top
@chapter Restrictions

@itemize @bullet
@item
Demacs does not support asyncronous process. Therefore you can not use
standard @file{compile.el}, @file{shell-mode.el} and etc. I think it's
too painful, so, customized @file{compiled.el} and @file{dired.el} are
provided. See lisp directory.

@item
@kbd{Ctrl-G} can't generate inturrupt signal. So you can't quit lisp-form
like this:

@example
(while t ())
@end example

@noindent
You shoud use the following form instead of it.

@example
(while (not (input-pending-p)) ())
@end example

@item
Demacs backup filename convension is following.

Base backup filename is made from following rule.

@example
foo      ->  foo.~
foo.c    ->  foo.c~
foo.tex  ->  foo.te~
@end example

Since DOS filename rule (8.3) is too tight, numeric backup such like
@file{foo.c.~2~} can not be used.

Similarly auto save filename is made from following rule.

@example
abcdefgh ->  #abcdefg.#
foo.c    ->  #foo.c#
foo.tex  ->  #foo.te#
@end example

Similarly, auto-save file name is created.

@item
If you don't configure XMS or VCPI driver and invoke Demacs with
@code{-keep} option, child process Demacs (or go32 application) will
@t{*}not@t{*} work correctly.@refill

@item
If you use VCPI manager and you have less extended memory than 300KB,
Demacs will hang up or being to work @t{*}VERY@t{*} slowly. Plese give
Demacs much memory if you met this.@refill
@end itemize


@node Compilation, Appendix, Restrictions, Top
@chapter Compilation

If you wish to recompile Demacs, 'GJ's GCC/G++' (GNU C compiler for DOS)
and original Emacs source are required.@refill

Diffs to the original Emacs source files are included in
@file{src\diffs.lzh}. Unpack the archive. The files in @file{diffs.lzh}
are not @code{diff} form but they are full source files, so you need not to
apply patch.  You may need to alter makefile.@refill

Then run @samp{make -f xmakefile all} command. Make generates
@code{temacs} and @code{demacs}.  Caution! just @code{make} will destroy
@file{xmakefile}. I prepared @file{xmake.bak} for you.@refill

Note that all modifications are embeded by @code{#ifdef MSDOS} and
@code{#endif MSDOS}, so you may follow my modifications easily. Machine
specified codes are embedded in similar flags.

@example
    *.h *.c (C source files)
       |
 ------| make
 |     | 
 |     V
 |   temacs (plain demacs) and lisp/*.el (lisp source files)
 |     | 
 |     | go32 temacs -batch -l loadup.el dump
 |---->|
       V
     xemacs
@end example

Provided @file{bin/temacs} enables you to reconfigure demacs without
recompilation. After editing lisp sources which are loaded up by
@file{loadup.el}, run @samp{go32 temacs -batch -l loadup.el dump}.


@node Appendix, TerminalFeatures, Compilation, Top

@menu
* TerminalFeatures::  Terminal Specified Features.
* DiffsFromUNIX::     Differences from UNIX Version.
* Index::             Index: Functions and Variables.
@end menu


@node TerminalFeatures, DiffsFromUNIX, Appendix, Appendix
@appendix Terminal Specified Features

If your machine is IBM PC or its compatibles, set environment variable
@var{TERM} with @code{ibmpc}, then you can use cursor keys and functon
keys.

The functions which is assigned to those keys are as follows:

@ifinfo
@display
Key        Function
----------+--------------------+-----------------
PgUp       scroll-up            Scroll text upward by page        
PgDn       scroll-down          Scroll text downward by page      
Insert     set-mark-command     Set mark current cursor position  
Del        delete-backward-char Delete previous character         
UpArrow    previous-line        Move cursor vertically up         
LeftArrow  backward-char        Move cursor left                  
RightArrow forward-char         Move cursor right                 
DownArrow  next-line            Move cursor vertically down       
Home       beginning-of-buffer  Move cursor beginning of buffer   
End        end-of-buffer        Move cursor end of buffer         
@end display
@end ifinfo

@tex
\vskip \baselineskip
\indent
\halign{\tt #\hfil&\qquad#&\tt #\hfil&\qquad#&\rm #\hfil\cr
PgUp       &&scroll-up            &&Scroll text upward by page\cr
PgDn       &&scroll-down          &&Scroll text downward by page\cr
Insert     &&set-mark-command     &&Set mark current cursor position\cr
Del        &&delete-backward      &&Delete previous character\cr
UpArrow    &&previous-line        &&Move cursor vertically up\cr
LeftArrow  &&backward-char        &&Move cursor left\cr
RightArrow &&forward-char         &&Move cursor right\cr
DownArrow  &&next-line            &&Move cursor vertically down\cr
Home       &&beginning-of-buffer  &&Move cursor beginning of buffer\cr
End        &&end-of-buffer        &&Move cursor end of buffer\cr
}
@end tex

@ifinfo
@display
F1         help-for-help        Show help for help
Shift-F1   describe-mode        Show help of current mode
Ctrl-F1    describe-key         Show help of key

F2         other-window         Select different window
Shift-F2   switch-to-buffer     Select buffer by name
Ctrl-F2    list-buffers         Display a list of buffers

F3         delete-other-windows Make current window fill the screen
Shift-F3   delete-window        Remove current window
Ctrl-F3    scroll-other-window  Scroll text of other window upward by page

F4         split-window-vertically      Split current window vertically
Shift-F4   enlarge-window               Make current window bigger
Ctrl-F4    shrink-window                Make current window smaller

F5         split-window-horizontally    Split current window horizontally
Shift-F5   enlarge-window-horizontally  Make current window wider
Ctrl-F5    shrink-window-horizontally   Make current window narrower

F6         copy-region          Copy from mark till cursor to yank-buffer
Shift-F6   kill-region          Cut from mark till cursor to yank-buffer
Ctrl-F6    yank                 Paste yank-buffer to current cursor position

F7         goto-line            Goto line
Shift-F7   isearch-forward      Do incremental search forward
Ctrl-F7    query-replace        Do query replace

F8         find-file            Edit file
Shift-F8   save-buffer          Save current buffer
Ctrl-F8    write-file           Write current buffer into file

F9         dired                        'Edit' directory
Shift-F9   compile                      Compile the program
Ctrl-F9    repeat-complex-command       Edit and re-evaluate last command

F10        shell-command                Execute command
Shift-F10  suspend-emacs                Call DOS command interpreter
Ctrl-F10   save-buffers-kill-emacs      Quit Emacs
@end display
@end ifinfo

@tex
\vskip \baselineskip
\indent
\halign{\tt #\hfil&\quad#&\tt #\hfil&\quad#&\rm #\hfil\cr
F1	   &&help-for-help		&&Show help for help\cr
Shift-F1   &&describe-mode		&&Show help of current mode\cr
Ctrl-F1	   &&describe-key		&&Show help of key\cr
\cr
F2	   &&other-window		&&Select different window\cr
Shift-F2   &&switch-to-buffer		&&Select buffer by name\cr
Ctrl-F2	   &&list-buffers		&&Display a list of buffers\cr
\cr
F3	   &&delete-other-windows	&&Make current window fill the screen\cr
Shift-F3   &&delete-window		&&Remove current window\cr
Ctrl-F3	   &&scroll-other-window	&&Scroll text of other window upward by page\cr
\cr
F4	   &&split-window-vertically	&&Split current window vertically\cr
Shift-F4   &&enlarge-window		&&Make current window bigger\cr
Ctrl-F4	   &&shrink-window		&&Make current window smaller\cr
\cr
F5	   &&split-window-horizontally	&&Split current window horizontally\cr
Shift-F5   &&enlarge-window-horizontally&&Make current window wider\cr
Ctrl-F5	   &&shrink-window-horizontally	&&Make current window narrower\cr
\cr
F6	   &&copy-region	&&Copy from mark till cursor to yank-buffer\cr
Shift-F6   &&kill-region	&&Cut from mark till cursor to yank-buffer\cr
Ctrl-F6	   &&yank		&&Paste yank-buffer to current cursor position\cr
\cr
F7	   &&goto-line		&&Goto line\cr
Shift-F7   &&isearch-forward	&&Do incremental search forward\cr
Ctrl-F7	   &&query-replace	&&Do query replace\cr
\cr
F8	   &&find-file		&&Edit file\cr
Shift-F8   &&save-buffer	&&Save current buffer\cr
Ctrl-F8	   &&write-file		&&Write current buffer into file\cr
\cr
F9	   &&dired			&&'Edit' directory\cr
Shift-F9   &&compile			&&Compile the program\cr
Ctrl-F9	   &&repeat-complex-command	&&Edit and re-evaluate last command\cr
\cr
F10	   &&shell-command		&&Execute command\cr
Shift-F10  &&suspend-emacs		&&Call DOS command interpreter\cr
Ctrl-F10   &&save-buffers-kill-emacs	&&Quit Emacs\cr
}
\vskip \baselineskip
@end tex

@noindent
Summary:

@ifinfo
 | F.1 |  F.2  |  F.3  |  F.4   |  F.5   | F.6 |  F.7  | F.8  | F.9  | F.10   |
 |help | nextW | delOW |splitWH |splitWV |copy | goto  | read |dired |command |
s|mode |switchW| delW  |largeWH |largeWV |kill |search | save | make |suspend |
c|key  |buffers|scrlOW |shrnkWH |shrnkWV |yank |replace|write |repeat| quit   |
@end ifinfo

@tex
\vskip \baselineskip
\indent
\vbox{\offinterlineskip
\halign{
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil&
\vrule#&
\hfil\ \tt #\hfil\cr
 &&  F.1  &&  F.2  &&  F.3  &&  F.4  &&  F.5  &&  F.6  && F.7 &&  F.8  && F.9 && F.10   \cr
\omit&height2pt&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit\cr
\noalign{\hrule}
\omit&height2pt&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit&&\omit\cr
  && help  &&  nextW  &&  delOW  && splitWH  && splitWV  && copy  &&  goto   &&  read  && dired  && command  \cr
s && mode  && switchW &&  delW   && largeWH  && largeWV  && kill  && search  &&  save  &&  make  && suspend  \cr
c && key   && buffers && scrlOW  && shrnkWH  && shrnkWV  && yank  && replace && write  && repeat &&  quit    \cr
}}
\vskip \baselineskip
@end tex


If you want to append and/or change the function, use lisp variable
@code{term-setup-hook}. See @file{lisp/term/ibmpc.el}.@refill

If you don't need cursor and function keys, append next line to
@file{$@var{HOME}/_emacs}:@refill

@example
(setq term-file-prefix nil)
@end example


@node DiffsFromUNIX, OmittedFunctions, TerminalFeatures, Appendix
@appendix Differences from UNIX Version

@menu
* OmittedFunctions:: Not implemented Functions to Demacs.
* OmittedVariables:: Not implemented Variables to Demacs.
* AddedFunctions::   Added Functions to Demacs.
* AddedVariables::   Added Variables to Demacs.
@end menu


@node OmittedFunctions, OmittedVariables, DiffsFromUNIX, DiffsFromUNIX
@unnumberedsec List of Functions which are not inplemented to Demacs

@example
accept-process-output                   process-list
continue-process                        process-mark
delete-process                          process-name
file-locked-p                           process-send-eof
get-buffer-process                      process-send-region
get-process                             process-send-string
interrupt-process                       process-sentinel
kill-process                            process-status
list-processes                          processp
make-symbolic-link                      quit-process
open-network-stream                     set-process-buffer
process-buffer                          set-process-filter
process-command                         set-process-kanji-code
process-exit-status                     set-process-sentinel
process-filter                          start-process
process-id                              stop-process
process-kanji-code                      waiting-for-user-input-p
process-kill-without-query
@end example


@node OmittedVariables, AddedFunctions, OmittedFunctions, DiffsFromUNIX
@unnumberedsec List of Variables which are not inplemented to Demacs

@example
delete-exited-processes                 process-connection-type
@end example


@node AddedFunctions, AddedVariables, OmittedVariables, DiffsFromUNIX
@unnumberedsec List of Functions added to Demacs

@example
bdos                                    intdos
define-file-name-file-type              invoke-find-file-type
demacs-version                          make-register
disk-free-space                         register-value
disk-total-space                        set-cursor-mode
file-type-internal                      set-default-file-type
file-type-p                             set-file-type
find-file-not-found-set-file-type       set-keyclick
find-file-type-from-file-name           set-register-value
get-cursor-mode                         set-screen-mode
get-screen-mode                         toggle-pass8-mode
int86
@end example


@node AddedVariables, Index, AddedFunctions, DiffsFromUNIX
@unnumberedsec List of Functions added to Demacs

@example
default-file-type                       file-type
demacs-version                          file-type-alist
demacs-version-date                     find-file-type
dos-inhibit-setdisk                     pass8-mode
dos-machine-type                        register-name-by-byte-alist
file-name-file-type-alist               register-name-by-word-alist
@end example


@node Index, , AddedVariables, Appendix
@appendix Index
@unnumberedsec Functions
@printindex fn
@unnumberedsec Variables
@printindex vr

@contents
@bye
